CUDA-accelerated genetic feedforward-ANN training for data mining

### 2010-07-09 doc/HPCS-proceedings-paper.tex
Add references and further explanation on method of standardization of feature values
git-svn-id: http://svn.scesoc.ca/evolution/trunk@125 fee98b4c-d1c3-4d34-9525-a468159e4a48

To avoid this issue, we standardize the data before training and before classification. For each feature individually, we estimate the 10\textsuperscript{th} and 90\textsuperscript{th} percentiles of the feature values. We then shift and scale the values such that 80\% of the resulting values are in the range $(-1, 1)$. This standardization method was chosen for its robustness against outliers. After standardization, the range of inputs to \texttt{expf} is narrow enough to always produce a valid FP value at the output. The standardization does not affect the classification itself because there are scaling and shifting parameters within the ANN (input to hidden edge weights and hidden center parameters, respectively) which are already part of the search space.

>>>>

To avoid this issue, we standardize each feature individually before training and before classification. One method of standardization is to use the sample mean and sample standard deviation as shifting and scaling parameters (referred to as "z-score normalization" in \cite{ZscoreNormalization}). However, sample standard deviation is not robust against outliers, while interquartile range (IQR) does tolerate outliers \cite{RobustRegression}. We use a variant of IQR: we estimate the range between the 10\textsuperscript{th} and 90\textsuperscript{th} percentiles and use that as the scaling parameter. After shifting and scaling, 80\% of the resulting feature values are in the range $(-1, 1)$. This range of inputs to \texttt{expf} is narrow enough to always produce a valid FP value at the output. The standardization does not affect the classification itself because there are scaling and shifting parameters within the ANN (input to hidden edge weights and hidden center parameters, respectively) which are already part of the search space.
